---
title: "MCMCglmm Significance Testing of Body Condition Heritability and Repeatability in Brown Treesnakes (*Boiga irregularis*) - 16 November 2020 update"
output:
  html_notebook: default
  word_document: default
  pdf_document: default
---

#Overview

This R notebook details significance testing of heritability and repeatability for body condition, a trait related to Brown Treesnake control and reproduction. Refer to Levine *et al*. (2019) for details regarding this data set. The pedigree includes 73 dummy individuals.


Based off of sensitivity analysis of priors, the priors used assume phenotypic variation is split evenly between additive genetic, individual permanent environment, and residual variance components with a low degree of belief (n = 0.2).


The prior was specified with the code: 
list(G = list(G1 = list(V = matrix(p.var/3), nu = 0.2), G2 = list(V = matrix(p.var/3), nu = 0.2)), R = list(V = matrix(p.var/3, nu = 0.2).


p.var = phenotypic variance of the trait as estimated from the data.


Significance of heritability was tested by comparing the deviance information criteria for models containing the "animal" term versus those of models without the "animal" term. A lower DIC is indicative of the preferred model, while DIC < 10 are strong evidence of a preferred model (Martin and Festa-Bianchet 2012).


Differently, significance of repeatability for each trait was testing by comparing the deviance information criteria for models containing the "individual" term (but not the "animal" term; note: additive genetic variation is a part of individual variation) versus models without the "individual" term.


For reduced models, the prior is adjusted to split phenotypic variation evenly among the remaining variance commponents.


#1. Significance Testing -- Heritability

##Body Condition (BCI) Heritability
```{r}
#load package
library(MCMCglmm)

#load model with "animal" term for BCI
load("BCI_1_dummy_SVLfilt.RData")

#view model to confirm it's correct
summary(model_BCI1)

#view priors to confirm their correct
prior_BCI1

#extract posterior mode of Va
Va <- posterior.mode(model_BCI1$VCV[, "animal"])

#extract HPD intervals for Va
x <- HPDinterval(model_BCI1$VCV[, "animal"])

#cat Va estimate with 95% HPD
cat("The Va estimate for BCI is: ", round(Va, digits = 3), "(", round(x[,1], digits = 3), "-", round(x[,2], digits = 3), ")", "\n")

#extract posterior mode of Vi
Vi <- posterior.mode(model_BCI1$VCV[, "ID"])

#extract HPD intervals for Va
x <- HPDinterval(model_BCI1$VCV[, "ID"])

#cat Va estimate with 95% HPD
cat("The Vi estimate for BCI is: ", round(Vi, digits = 3), "(", round(x[,1], digits = 3), "-", round(x[,2], digits = 3), ")", "\n")

#extract posterior mode of Vr
Vr <- posterior.mode(model_BCI1$VCV[, "units"])

#extract HPD intervals for Va
x <- HPDinterval(model_BCI1$VCV[, "units"])

#cat Va estimate with 95% HPD
cat("The Vr estimate for BCI is: ", round(Vr, digits = 3), "(", round(x[,1], digits = 3), "-", round(x[,2], digits = 3), ")", "\n")

#calculate Vp
Vp <- model_BCI1$VCV[, "animal"] + model_BCI1$VCV[, "ID"] + model_BCI1$VCV[, "units"]

#HPD intervals for Vp
x <- HPDinterval(Vp)

#cat Vp estimate with 95% HPD
cat("The Vp estimate for BCI is: ", round(posterior.mode(Vp), digits = 3), "(", round(x[,1], digits = 3), "-", round(x[,2], digits = 3), ")", "\n")

#calculate heritability
herit <- model_BCI1$VCV[, "animal"]/(model_BCI1$VCV[, "animal"] + model_BCI1$VCV[, "ID"] + model_BCI1$VCV[, "units"] )

#HPD intervals for heritability
x <- HPDinterval(herit)

#cat heritability estimate with 95% HPD
cat("The heritability estimate for BCI is: ", round(posterior.mode(herit), digits = 3), "(", round(x[,1], digits = 3), "-", round(x[,2], digits = 3), ")", "\n")

#extract DIC from model with "animal" term
DIC_full <- model_BCI1$DIC

#load model without "animal" term for BCI 
load("BCI_1_SVLfilt_noVa.RData") #note: over-writing earlier BCI model

#view model to confirm it's correct
summary(model_BCI1)

#view priors to confirm they're correct
prior_BCI1

#extract DIC for model without "animal" term for BCI
DIC_reduced <- model_BCI1$DIC

#cat statement about the DICs of the models
cat("The DIC of the model with the animal term is: ", round(DIC_full, digits = 3), "\n", "The DIC of the model without the animal term is: ", round(DIC_reduced, digits = 3), "\n\n\n")

#calculate difference between DICs
diff <- DIC_reduced - DIC_full

#test whether heritability is significant
if(diff >= 10) {
  cat("BCI model with animal term has smaller DIC: ", DIC_full < DIC_reduced, "\n", "BCI model with animal term has SIGNIFICANTLY smaller DIC:", diff >= 10, "\n", "BCI has significant heritability.\n") 
} else {
  cat("BCI model with animal term has smaller DIC: ", DIC_full < DIC_reduced, "\n", "BCI model with animal term has SIGNIFICANTLY smaller DIC:", diff >= 10, "\n", "BCI does not have significant heritability.")
}

```

#2. Significance Testing - Repeatability

##Body Condition (BCI) Repeatability
```{r}
#clear memory
rm(list=ls())

#load model with only "individual" term
load("BCI_1_SVLfilt_noVa.RData")

#view model to confirm it's correct
summary(model_BCI1)

#view priors to confirm their correct
prior_BCI1

#calculate repeatability
repeatability <- model_BCI1$VCV[, "ID"]/(model_BCI1$VCV[, "ID"] + model_BCI1$VCV[, "units"] )

#HPD intervals for repeatability
x <- HPDinterval(repeatability)

#cat statement about repeatability with 95% HPD
cat("The repeatability estimate for BCI is: ", round(posterior.mode(repeatability), digits = 3), "(", round(x[,1], digits = 3), "-", round(x[,2], digits = 3), ")", "\n")

#extract DIC from model with "ID" term
DIC_full <- model_BCI1$DIC

#load model without "ID" term for BCI 
load("BCI_1_SVLfilt_noVa_noI.RData") #note: over-writing earlier BCI model

#view model to confirm it's correct
summary(model_BCI1)

#view priors to confirm they're correct
prior_BCI1

#extract DIC for model without "ID" term for BCI
DIC_reduced <- model_BCI1$DIC

#cat statement about the DICs of the models
cat("The DIC of the model with the ID term is: ", round(DIC_full, digits = 3), "\n", "The DIC of the model without the ID term is: ", round(DIC_reduced, digits = 3), "\n\n\n")

#calculate difference between DICs
diff <- DIC_reduced - DIC_full

#test whether repeatability is significant
if(diff >= 10) {
  cat("BCI model with ID term has smaller DIC: ", DIC_full < DIC_reduced, "\n", "BCI model with ID term has SIGNIFICANTLY smaller DIC:", diff >= 10, "\n", "BCI has significant repeatability\n") 
} else {
  cat("BCI model with ID term has smaller DIC: ", DIC_full < DIC_reduced, "\n", "BCI model with ID term has SIGNIFICANTLY smaller DIC:", diff >= 10, "\n", "BCI does not have significant repeatability")
}

```

