---
title: "MCMCglmm Diagnostics -- Brown Treesnake (*Boiga irregularis*) Heritability"
output: html_notebook
---

# Overview
This R notebook details the testing of two different priors in variance component estimation for 3 traits important to Brown Treesnake control and reproduction: body condition (BCI), snout-vent length (SVL), and trappability. Refer to Levine *et al*. (2019) for details regarding this data set. The pedigree includes 73 dummy individuals.

The two different sets of priors tested are:

**Prior 1**: Phenotypic variation split evenly between additive genetic, individual permanent environment, and residual variance components with a low degree of belief (n = 0.2).

Prior 1 was specified with the code: list(G = list(G1 = list(V =           matrix(p.var/3), nu = 0.2), G2 = list(V = matrix(p.var/3), nu = 0.2)), R = list(V = matrix(p.var/3, nu = 0.2) 

**Prior 2**: 95% of phenotypic variance is due to the additive genetic effect, with the remaining 5% of phenotypic variance due evenly to individual permanent environment (2.5%) and residual (2.5%) effects, with a low degree of belief (n = 0.2).

Prior 2 was specified with the code: list(G = list(G1 = list(V = matrix(p.var x 0.95), nu = 0.2), G2 = list(V = matrix(p.var x 0.025), nu = 0.2), R = list(V = matrix(p.var x 0.025), nu = 0.2)

For both priors, p.var = phenotypic variance of the trait as estimated from the data.

**Note**: The priors were tested for each trait, denoted by 1 or 2. For example, BCI1 is the model for BCI tested with prior 1, BCI2 is the model for BCI tested with prior 2. 


# MCMCglmm Diagnostics for BCI--

## BCI1
```{r}
#load the MCMCglmm package
library(MCMCglmm)

#load R output
load("BCI_1_dummy_SVLfilt.RData")

#set the dimensions for the graphics
par(mar=c(1,1,1,1))

#view model
summary(model_BCI1)

#view priors for model
prior_BCI1

#view trace plots for fixed effects
plot(model_BCI1$Sol)

#view trace plots for random effects
plot(model_BCI1$VCV)

#view autocorrelation for fixed effects
autocorr.diag(model_BCI1$Sol)

#view autocorrelation for random effects
autocorr.diag(model_BCI1$VCV)

#view effective size for fixed effects
effectiveSize(model_BCI1$Sol)

#view effective size for random effects
effectiveSize(model_BCI1$VCV)

#test for convergence of fixed effect
heidel.diag(model_BCI1$Sol)

#test for convergence of random effect
heidel.diag(model_BCI1$VCV)

#view posterior mode of additive genetic variance
cat("Posterior Mode of Additive Genetic Variance for BCI1: ", round(posterior.mode(model_BCI1$VCV[, "animal"]), digits = 3))

#view posterior mode of individual
cat("Posterior Mode of Individual Variance for BCI1: ", round(posterior.mode(model_BCI1$VCV[, "ID"]), digits = 3))

#calculate heritability
herit <- model_BCI1$VCV[, "animal"]/(model_BCI1$VCV[, "animal"] + model_BCI1$VCV[, "ID"] + model_BCI1$VCV[, "units"] )

#view posterior mode of heritability
cat("Posterior Mode of Heritability for BCI1: ",round(posterior.mode(herit), digits = 3) )

#calculate repeatability
repeatability <- (model_BCI1$VCV[, "animal"] + model_BCI1$VCV[, "ID"])/(model_BCI1$VCV[, "animal"] + model_BCI1$VCV[, "ID"] + model_BCI1$VCV[, "units"])

#view posterior mode of repeatability
cat("Posterior Mode of Repeatability for BCI1: ",round(posterior.mode(repeatability), digits = 3) )
```
## BCI2
```{r}
#clear memory
rm(list=ls())

#load R output
load("BCI_2_dummy_SVLfilt.RData")

#set the dimensions for the graphics
par(mar=c(1,1,1,1))

#view model
summary(model_BCI2)

#view priors for model
prior_BCI2

#view trace plots for fixed effects
plot(model_BCI2$Sol)

#view trace plots for random effects
plot(model_BCI2$VCV)

#view autocorrelation for fixed effects
autocorr.diag(model_BCI2$Sol)

#view autocorrelation for random effects
autocorr.diag(model_BCI2$VCV)

#view effective size for fixed effects
effectiveSize(model_BCI2$Sol)

#view effective size for random effects
effectiveSize(model_BCI2$VCV)

#test for convergence of fixed effect
heidel.diag(model_BCI2$Sol)

#test for convergence of random effect
heidel.diag(model_BCI2$VCV)

#view posterior mode of additive genetic variance
cat("Posterior Mode of Additive Genetic Variance for BCI2: ", round(posterior.mode(model_BCI2$VCV[, "animal"]), digits = 3))

#view posterior mode of individual
cat("Posterior Mode of Individual Variance for BCI2: ", round(posterior.mode(model_BCI2$VCV[, "ID"]), digits = 3))

#calculate heritability
herit <- model_BCI2$VCV[, "animal"]/(model_BCI2$VCV[, "animal"] + model_BCI2$VCV[, "ID"] + model_BCI2$VCV[, "units"] )

#view posterior mode of heritability
cat("Posterior Mode of Heritability for BCI2: ",round(posterior.mode(herit), digits = 3) )

#calculate repeatability
repeatability <- (model_BCI2$VCV[, "animal"] + model_BCI2$VCV[, "ID"])/(model_BCI2$VCV[, "animal"] + model_BCI2$VCV[, "ID"] + model_BCI2$VCV[, "units"])

#view posterior mode of repeatability
cat("Posterior Mode of Repeatability for BCI2: ",round(posterior.mode(repeatability), digits = 3) )
```


`




